### 2022-05-25 analyses/plotterScript.Rmd

---
title: "plottingScript"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(casebase) #core
library(keras) #core
source("src/functions.R") #core
library(data.table)#corecr
library(mltools)#corecr

library(survival)#compare
library(survminer)#compare
library(ggplot2)#compare
library(colorspace)#compare
library(DT)
library(splines)
library(reticulate)

library(simsurv)
library(flexsurv)
library(survivalmodels)
library(visreg)

#library(tidyverse)
library(riskRegression)
library(pseudo)

library(tensorflow)
library(ggrepel)

library(dplyr)
library(flextable)

library(cowplot)
library(matrixStats)
library(tableone)
library(gridExtra)
library(grid)
library(gtable)
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)
set.seed(1)
epo=2000
bsize=512
min_delta = 10^-7
patience = 10
iteration=50
e_sup=F
eval_firstSim<-F
numSamp<-2000

 pal <- function(col, border = "light gray", ...)  {
   n <- length(col)
   plot(0, 0, type="n", xlim = c(0, 1), ylim = c(0, 1),
   axes = FALSE, xlab = "", ylab = "", ...)
   rect(0:(n-1)/n, 0, 1:n/n, 1, col = col, border = border)
   }
statCalcSim<-function(metricSims){

toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
if(length(toRemove)>0){
metricSims<-metricSims[-c(toRemove),]
}
  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),means=rowMeans(as.matrix(metricSims[,-c(1,2)])))

  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
  rownames(ci)<-c('cilower95','ciupper95')
  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
  rownames(pi)<-c('pilower95','piupper95')

  metricSummary<-cbind(metricSummary,t(ci),t(pi))

  return(metricSummary)
}

plotPerformance<-function(metricSummary,xlab="Follow-up time",method='Metric',iterations='NA'){

  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','PMNN_Spline','PMNN_Poly','Optimal',
                                "dsm"),
                        color=qualitative_hcl(9,palette='Dark 3'),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                'DeepHit','PMNN_Spline','CBNN','Optimal',
                                "DSM"))
  allModels$color[1]="#bfc1c2"
  allModels$color[2]="#FFED11"#"#7C4C74"
  allModels$color[3]= "#80471C" #"#F9A602"
  allModels$color[5]="#FF00AB"
  allModels$color[8]="#000000"
  allModels$color[9]="#8077D5"
 allModels<- allModels[order(allModels$trueName),]
    i <- sapply(metricSummary, is.factor)
    metricSummary[i] <- lapply(metricSummary[i], as.character)
    metricSummary<-metricSummary[complete.cases(metricSummary), ]
    mainTitle<-""
    metricSummary$color<-NA
    for(i in 1:nrow(metricSummary)){
      modelPos<-which(metricSummary$model[i] ==allModels$model)
      metricSummary$model[i]<-allModels$trueName[modelPos]
      metricSummary$color[i]<-allModels$color[modelPos]

    }

   colorsToUse<- sort(which(allModels$trueName %in% metricSummary$model))

    if(method=="C-Index"){
    method=expression("c-ipcw")
    }   
 metricSummary<- metricSummary[-which(metricSummary$model=="PMNN_Spline"),]
  #mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
  colnames(metricSummary)[1]<-"Models"
  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=Models))+
    geom_line() +
    geom_ribbon(aes(ymin=cilower95,ymax=ciupper95,fill=Models),linetype=0,alpha=0.3) +
    ggtitle(mainTitle) + 
    xlab(xlab) +
    ylab(method)+
    theme_minimal()+
    scale_color_manual(values=allModels$color[colorsToUse])+
    scale_fill_manual(values=allModels$color[colorsToUse])
   # scale_color_manual(labels = metricSummary$model, values = metricSummary$color)

  return(a)
}

```

```{r plotsupport,eval=T}

#age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine
BrierIPAList<-readRDS('results/100supBrier.rds')
BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(support$d.time, na.rm = T)-min(support$d.time, na.rm = T)) + min(support$d.time, na.rm = T)
ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)

ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}

ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall

ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
data<-casebase::support
covs<-colnames(data)
catVars<-c('death','sex','dzgroup','dzclass','race','ca')

data[,-which( covs%in%catVars )]<-as.data.frame(sapply(data[,-which( covs%in%catVars )], function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))

colnames(data)[5]<-'time'
colnames(data)[2]<-'status'
ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/supportResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)
#catVars<-c('x4','x5','x6','x7','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
saveRDS(tab2,"figures/t1Support.rds")

pdf("figures/t1Support.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2)
dev.off()

rm(pt)
#min(support$d.time, na.rm = T)
#max(support$d.time, na.rm = T)

ipaSu<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierSu<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsSu<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100supcidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxSu<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
cidxSu

```

```{r plotMetabric,fig.height=6,fig.width=6}

BrierIPAList<-readRDS('results/100MetaBrier.rds')
source_python("src/getMetabricData.py")
metabric<-getMetabric()

colnames(metabric)<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
BrierIPAList[[1]]$times<-BrierIPAList[[1]]$times*(max(metabric$time, na.rm = T)-min(metabric$time, na.rm = T)) + min(metabric$time, na.rm = T)
ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

colnames(metabric)[ncol(metabric)-1]<-'time'
colnames(metabric)[ncol(metabric)]<-'status'
data<-metabric
#data<-as.data.frame(sapply(data, function(x) (x - min(x, na.rm = T)) / (max(x, na.rm = T) - min(x, na.rm=T))))
covs<-colnames(data)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=0.5)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/metabricResult.pdf")
plot_grid(pt,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))#, legend
dev.off()

myVars<-c('MK167','EGFR','PGR','ERBB2','Hormone treatment','radiotherapy','Chemotherapy','ER_positive','age at diagnosis','time','status')
catVars<-c('Hormone treatment','radiotherapy','Chemotherapy','ER_positive','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars)))
saveRDS(tab2,"figures/t1Metabric.rds")

pdf("figures/t1metabric.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2)
dev.off()
rm(pt)

ipaM<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierM<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsM<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

ipa
ibs
brier

cidxList<-readRDS('results/100Metacidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxM<-plotPerformance(cidxSummary,method='C-Index',iterations=iteration  )
cidxM
```

```{r plotsimple,eval=T}

BrierIPAList<-readRDS('results/100easyBrier.rds')

ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))

dat <- simsurv(dist = "exponential", lambdas = 1, betas = c(z1 = .1,z2=.1,z3=.1),
                  x = cov, tdefunction = "log", maxt = 1)

  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=.01)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa,
          brier,#+ theme(legend.position="none"),
          ibs,#+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/simpleResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)[-1]
catVars<-c('z1','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars,strata='z1')))
tab2<-tab2[,-c(3,4)]
tab2[2,1]<-"0 (0.0)"
tab2[2,2]<-"1000 (100.0)"
tab2[3,1]<-"0 (0.50)"
tab2[3,2]<-"10 (0.50)"
tab2[4,1]<-"8 (0.50)"
tab2[4,2]<-"-3 (0.50)"
tab2<-tab2[-c(5,6),]
tab2
saveRDS(tab2,"figures/covariatesTable.rds")

pdf("figures/t1Simple.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2[,-ncol(tab2)])
dev.off()

rm(pt)

ipaS<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierS<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsS<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100easycidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxS<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )

```

```{r complexPrep}

# Define a function returning the log cum hazard at time t
logcumhaz <- function(t, x, betas, knots) {

  # Obtain the basis terms for the spline-based log
  # cumulative hazard (evaluated at time t)
  basis <- flexsurv::basis(knots, log(t))

  # Evaluate the log cumulative hazard under the
  # Royston and Parmar specification
  res <-betas[["gamma0"]] * basis[[1]] + 
    betas[["gamma1"]] * basis[[2]] +
    betas[["gamma2"]] * basis[[3]] +
    betas[["gamma3"]] * basis[[4]] +
    betas[["gamma4"]] * basis[[5]] +
    betas[["z1"]] * x[["z1"]]+
    betas[["z2"]] * x[["z2"]]+
    betas[["z3"]] * x[["z3"]]+
    betas[["term1"]] * x[["z1"]]*x[["z2"]]*t+
    betas[["term2"]] * x[["z1"]]*x[["z3"]]+
    betas[["term3"]] * x[["z2"]]*(x[["z3"]])
  #
  # Return the log cumulative hazard at time t
  res
}

pyProcess<-function(predictions,times=times){
  # test<-t(data.frame(a=seq(0,100,length.out = 200),b=seq(0,100,length.out = 200),c=seq(0,100,length.out = 200)))
  #predictions[is.na(predictions)] <- 0
  interpolated <- apply(predictions, 1, function(x) approx(as.numeric(colnames(predictions)), y=x, method = "linear", xout=times)$y)

  class(interpolated)<-c("tunnel",class(interpolated))
  interpolated<-1-interpolated

  interpolated[is.na(interpolated)] <- 0
  return(t(interpolated))
}

url<-'app.txt'
ds <- read.table(url, sep = '',header = TRUE)

ds$years.at.risk=pmin(ds$agein80,ds$onset)

#cov <- data.frame(id = 1:2000, hormon = rbinom(2000, 1, 0.5))

#ds$years.at.risk<-ds$years.at.risk/max(ds$years.at.risk)
#true_mod <- flexsurv::flexsurvspline(Surv(years.at.risk, app) ~ 1, data = ds, k = 3) #appendicitis from hanley

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))
  brcancer$rectime<-brcancer$rectime/max(brcancer$rectime)
  true_mod <- flexsurv::flexsurvspline(Surv(rectime, censrec) ~ 1, data = brcancer, k = 3) #from flexsurv

  # Simulate the event times
  coefficients<-true_mod$coefficients
  coefficients<-c(coefficients,z1=1,z2=1,z3=1,term1=10,term2=2,term3=2)
  #names(coefficients)<-c("shape",names(coefficients[-1]))

```

```{r plotComplex}

BrierIPAList<-readRDS('results/100hardBrier.rds')

ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times)
ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
count=1
for(sim in BrierIPAList){
  ipaSims<-cbind(ipaSims, sim$IPA)
  brierSims<-cbind(brierSims,sim$Brier)
  ibsSims<-cbind(ibsSims,sim$IBS)
  for(mod in colnames(ibsTable)){

    ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

  }
  count<-count+1
}
ibsTable<-na.omit(ibsTable)
ibsOverall<-data.frame(mean=apply(ibsTable,2,mean),sd = apply(ibsTable,2,sd))
ibsOverall
ipaSummary<-statCalcSim(ipaSims)
brierSummary<-statCalcSim(brierSims)
ibsSummary<-statCalcSim(ibsSims)
ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

#end_time-start_time

  cov<-data.frame(id=1:numSamp,
                z1=c(rep(1,numSamp/2),rep(0,numSamp/2)),
                z2=c(rnorm(numSamp/2,10,0.5),rnorm(numSamp/2,0,0.5)),
                z3=c(rnorm(numSamp/2,-3,0.5),rnorm(numSamp/2,8,0.5)))

  dat <- simsurv(betas = coefficients, # "true" parameter values
                 x = cov,                   # covariate data for 200 individuals
                 knots = true_mod$knots,    # knot locations for splines
                 logcumhazard = logcumhaz,  # definition of log cum hazard
                 maxt = 1,               # no right-censoring
                 interval = c(0,10000000)) # interval for root finding
  colnames(dat)<-c("id","time","status")
  samp<-sample(seq(1,nrow(dat),by=1), floor(nrow(dat)*(0.1))) #random censoring
  dat$status[samp]<-0
  # Merge the simulated event times onto covariate data frame
  data <- merge(cov, dat)

ptPlot<-plot(popTime(data,time="time",event="status"),percentile_number=.01)#+  theme(axis.text=element_text(size=8))
#end_time-start_time

pt <- plot_grid(ptPlot,
          ipa+ theme(legend.position="none"),
          brier+ theme(legend.position="none"),
          ibs+ theme(legend.position="none"),
          labels = c('A', 'B','C','D'))
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipa  +
    guides(color = guide_legend(nrow = 1)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

pdf(file="figures/complexResult.pdf")
plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
dev.off()

myVars<-colnames(data)[-1]
catVars<-c('z1','status')

tab2 <- as.data.frame(print(CreateTableOne(vars = myVars, data = data, factorVars = catVars,strata='z1')))
saveRDS(tab2,"figures/t1Complex.rds")

pdf("figures/t1Complex.pdf", height = 50*nrow(tab2), width = 200*ncol(tab2))
grid.table(tab2[,-ncol(tab2)])
dev.off()
rm(pt)

ipaC<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
brierC<-plotPerformance(brierSummary,method="BS",iterations=iteration )
ibsC<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )

cidxList<-readRDS('results/100hardcidx.rds')

temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
cidxSims=data.frame(model=temp$model,times=temp$times)

count=1
for(sim in cidxList){
  cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

  count<-count+1
}

cidxSummary<-statCalcSim(cidxSims)
cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
cidxC

```

```{r megaplot,eval=T,fig.height=12,fig.width=12}

pt <- plot_grid(
          ipaS+ theme(legend.position="none")+ ggtitle("Simple simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.5,0.1)),
          ipaC+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "",y="")+coord_cartesian(ylim=c(-.6,1)),
          ipaSu+ theme(legend.position="none")+ ggtitle("SUPPORT")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.20,.1)),
          ipaM+ theme(legend.position="none")+ ggtitle("METABRIC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.8,.25)),
          #brierS+ theme(legend.position="none")+ labs(x = ""),
          #brierC+ theme(legend.position="none")+ labs(x = "",y=""),
          #brierSu+ theme(legend.position="none")+ labs(x = "",y=""),
          #brierM+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsS+ theme(legend.position="none")+ labs(x = ""),
          #ibsC+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsSu+ theme(legend.position="none")+ labs(x = "",y=""),
          #ibsM+ theme(legend.position="none")+ labs(x = "",y=""), 
          cidxS+ theme(legend.position="none")+coord_cartesian(ylim=c(.45,.6)),
          cidxC+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.3,.9)),
          cidxSu+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.48,.6)),
          cidxM+ theme(legend.position="none")+ labs(y="")+coord_cartesian(ylim=c(0.3,.65)),
          #labels = c('A', 'E','I','M',"B",'F','J','N','C','G','K','O','D','H','L','P'))
          labels = c('A',"C","E","G",'B','D','F','H'),
          ncol=4)
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipaS  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))

pdf(file="figures/megaPlot.pdf",width=12,height=6)
g
dev.off()

saveRDS(g,"figures/megaPlot.rds")

```

```{r megaplotSupplemental,fig.height=12,fig.width=12}
pt <- plot_grid(
          #ipaS+ theme(legend.position="none")+ ggtitle("Simple simulation")+ labs(x = ""),
          #ipaC+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "",y=""),
          #ipaSu+ theme(legend.position="none")+ ggtitle("SUPPORT")+ labs(x = "",y=""),
          #ipaM+ theme(legend.position="none")+ ggtitle("METABRIC")+ labs(x = "",y=""),
          brierS+ theme(legend.position="none")+ labs(x = "")+coord_cartesian(ylim=c(0,.4))+ ggtitle("Simple simulation"),
          brierC+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3))+ ggtitle("Complex simulation"),
          brierSu+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0.1,.3))+ ggtitle("SUPPORT"),
          brierM+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.4))+ ggtitle("METABRIC"),
          ibsS+ theme(legend.position="none")+ labs(x = "")+coord_cartesian(ylim=c(0,.3)),
          ibsC+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)),
          ibsSu+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)),
          ibsM+ theme(legend.position="none")+ labs(x = "",y="")+coord_cartesian(ylim=c(0,.3)), 
          #cidxS+ theme(legend.position="none"),
          #cidxC+ theme(legend.position="none")+ labs(y=""),
          #cidxSu+ theme(legend.position="none")+ labs(y=""),
          #cidxM+ theme(legend.position="none")+ labs(y=""),
          #labels = c('A', 'E','I','M',"B",'F','J','N','C','G','K','O','D','H','L','P'))
          labels = c('A',"C","E","G",'B','D','F','H'),
          ncol=4)
#https://wilkelab.org/cowplot/articles/shared_legends.html
legend <- get_legend(
  # create some space to the left of the legend
  ipaS  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
   theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))

pdf(file="figures/megaPlotSupplement.pdf",width=12,height=6)
g
dev.off()

saveRDS(g,"figures/megaPlotSupplement.rds")

```

```{r flextableTablePrep,fig.height=46,fig.width=26}

edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
  newgrob <- editGrob(table$grobs[id][[1]], ...)
  table$grobs[id][[1]] <- newgrob
  }
  table
}

#SUMMARY DATA ORDER, IPA BS IBS CIDX
makeTable<-function(summaryData,study="study"){
   allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','PMNN_Spline','PMNN_Poly','Optimal',
                                "dsm"),
                        color=rainbow_hcl(9, start = 30, end = 300),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                'DeepHit','PMNN_Spline','CBNN','Optimal',
                                "DSM"))

  wantedScores<-c(0.25,0.5,0.75,1)

  tableResults<-as.data.frame(matrix(NA,nrow=nrow(allModels)+2,ncol=16))
  tableResults[1,]<-c("IPA","IPA","IPA","IPA",
                      "BS","BS","BS","BS",
                      "IBS","IBS","IBS","IBS",
                      "C-Index","C-Index","C-Index","C-Index"
                      )
  tableResults[2,]<-c(0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1)*100
  indexer<-1
  rownames(tableResults)<-c(study,"% of time",allModels$trueName)
  for (i in 1:length(summaryData)){ #IPA BRIER IBS CIDX is indexed by i

    tempPos<-which(rownames(tableResults) %in% unique(summaryData[[i]]$Models)  )
    tempModels<-rownames(tableResults)[ tempPos]
    for(j in 1:length(tempModels)){ #indexes model
      offset<-(i-1)*4
      indexer=1+offset
      for(k in 1:length(wantedScores)){ #indexes .25 .5 .75 1.0
        tempTime<-unique(summaryData[[i]]$times)[(floor(length(unique(summaryData[[i]]$times))*wantedScores[k]))]
        tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        if(length(round(summaryData[[i]]$means[tempResRow],2))<1){
          tableResults[tempPos[j],indexer]<-"-0.23 (-0.25,-0.23)" #note that this is grabbed directly from the last estimate of Cox, as its the only time we don't get a value. So we extrapolate from the last available survival time.
        }
        else{tableResults[tempPos[j],indexer]<-paste(round(summaryData[[i]]$means[tempResRow],2),
              " (",round(summaryData[[i]]$cilower95[tempResRow],2),
              ",",round(summaryData[[i]]$ciupper95[tempResRow],2),")",sep="")
        }
        indexer<-indexer+1

      }
    }

  }

  return(tableResults)
}

cleantables<-function(tab){
  toRemove<-which(rowSums(is.na(tab))==ncol(tab))
      tab[is.na(tab)]<-"-"
  if(length(toRemove)>0){

    return(tab[-toRemove,])
  }
  return(tab)
}

quickEdit<-function(grob,rows,columns){
  return(edit_cell(grob, rows, columns, "core-fg", 
               gp=gpar(fontsize=12, fontface="bold")))
}

addBorders<-function(grob){
grob$layout[1:2 , c("l", "r")] <- list(c(5, 9),c(2,6))
grob<- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 2, b = nrow(grob), l = 1, r = ncol(grob))
grob <- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 1, l = 1, r = ncol(grob))
grob <- gtable_add_grob(grob,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 1, l = 2,r=5,b= nrow(grob))
return(grob)
}

flexMaker<-function(tab,boldPicks=NA,title){
  boldPicks$col=boldPicks$col+1 #makes finding the correct ones easier.
bord<-fp_border_default(color="black",style="solid")
bord2<-fp_border_default(color="black",style="solid",width = 2)
tab<-cbind(rownames(tab),tab)
ft<-delete_part(flextable(tab[-c(1,2),],cwidth=5,cheight=1), part = "header")

ft <- add_header_row(x = ft, values = c(title,"IPA", "c-ipcw"),
  colwidths = c(1,4,4))
#ft<-bold(ft, bold = TRUE, part = "header")

#ft$header$styles$text$bold$data
ft <- add_header_row(x = ft,top=F, values = c("Method", 25,50,75,100,25,50,75,100),
  colwidths = c(1,1,1,1,1,1,1,1,1))
ft$header$styles$text$bold$data[1,1]<-T
#ft <- add_header_row(
#  x = ft, values = c(title),
#  colwidths = c(9))
ft <- align(ft, i = c(1,2), part = "header", align = "center")
ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
ft<-hline(ft,i=c(1,2),border=bord2,part="header")
#ft<-hline(ft,i=c(1),border=bord,part="body")
ft<-fontsize(ft, i = NULL, j = NULL, size =7.5, part = "all")
ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
ft <- align(ft, i = NULL, part = "body", align = "center")
ft$body$styles$text$font.size$data[,1]<-10

if(is.data.frame(boldPicks)){
  for(i in 1:nrow(boldPicks)){
    ft$body$styles$text$bold$data[boldPicks$row[i],boldPicks$col[i]]<-T
  }
}
return(rasterGrob(as_raster(autofit(ft))))
}

```

```{r flexTableMain,fig.height=46,fig.width=26}

sTab<-makeTable(list(ipaS$data,brierS$data,ibsS$data,cidxS$data),study="Simple")[-3,-c(5:12)]
#[-3,-c(1,2,3,4,13,14,15,16)] for the supplemental
cTab<-makeTable(list(ipaC$data,brierC$data,ibsC$data,cidxC$data),study="Complex")[-3,-c(5:12)]
suTab<-makeTable(list(ipaSu$data,brierSu$data,ibsSu$data,cidxSu$data),study="SUPPORT")[-3,-c(5:12)]
mTab<-makeTable(list(ipaM$data,brierM$data,ibsM$data,cidxM$data),study="METABRIC")[-3,-c(5:12)]
sTab<-cleantables(sTab)
cTab<-cleantables(cTab)
suTab<-cleantables(suTab)
mTab<-cleantables(mTab)
tempNames<-rownames(mTab)
mTab<-as.data.frame(lapply(mTab, function(x){gsub("0 ","0.00 ", x,fixed=T)}))
mTab<-as.data.frame(lapply(mTab, function(x){gsub('(0,',"(0.00,", x,fixed=T)}))
mTab<-as.data.frame(lapply(mTab, function(x) gsub(",0)",",0.00)", x,fixed=T)))
rownames(mTab)<-tempNames
sPicks<-data.frame(row=c(1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6,1,2,6),
                   col=c(1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8,1,2,3,4,5,6,7,8))
cPicks<-data.frame(row=c(6,6,6,6,5,5,5,2),
                   col=c(1,2,3,4,5,6,7,8))
suPicks<-data.frame(row=c(1,2,1,2,5,1,2,3,1,2,5,1,2,5,1,2,5,1,2,5),
                    col=c(1,1,2,2,2,3,3,4,5,5,5,6,6,6,7,7,7,8,8,8))
mPicks<-data.frame(row=c(2,4,5,1,2,1,2,4,5,4,4,5),
                   col=c(1,1,1,2,2,3,3,4,5,6,7,8))

sFlex<-flexMaker(tab=sTab,boldPicks = sPicks,title="A: Simple")
suFlex<-flexMaker(tab=suTab,boldPicks = suPicks,title="C: SUPPORT")
cFlex<-flexMaker(tab=cTab,boldPicks = cPicks,title="B: Complex")
mFlex<-flexMaker(tab=mTab,boldPicks = mPicks,title="D: METABRIC")

pdf(file="figures/megaTable.pdf",width=18,height=19.5)
cowplot::plot_grid(sFlex,cFlex, suFlex,mFlex, nrow = 4, ncol = 1)
dev.off()

```

```{r getMeans,fig.height=8,fig.width=16}

meanAndSort<-function(scores){
  meanScores<-aggregate(means~Models, data=scores, mean)
  return(tableGrob(meanScores[order(meanScores$means,decreasing=T),]))
}

simpleIPAMean<-meanAndSort(ipaS$data)
simplecidxMean<-meanAndSort(cidxS$data)

supportIPAMean<-meanAndSort(ipaSu$data)
supportcidxMean<-meanAndSort(cidxSu$data)

complexIPAMean<-meanAndSort(ipaC$data)
complexcidxMean<-meanAndSort(cidxC$data)

metaIPAMean<-meanAndSort(ipaM$data)
metacidxMean<-meanAndSort(cidxM$data)

pdf(file="figures/averageScores.pdf",width=16,height=8)
grid.arrange(simpleIPAMean,complexIPAMean,supportIPAMean,metaIPAMean,simplecidxMean,complexcidxMean,supportcidxMean,metacidxMean,nrow=2,ncol=4,top = "simple,complex,support,metabric")
dev.off()

```