### 2022-05-25 analyses/proofOfConceptWeights.Rmd

---
title: "realData"
output: html_document
---

```{r,eval=T}

library(reticulate)
```

```{python,eval=T}

import os
os.environ["CUDA_DEVICE_ORDER"] = "PCI_BUS_ID" 
os.environ["CUDA_VISIBLE_DEVICES"] = ""
#os.environ["OM_NUM_THREADS"]="2"

```

```{r setup, include=T}
set.seed(1)
knitr::opts_chunk$set(echo = TRUE)
source("src/packages.R")
source("src/functions.R")
source_python("src/cIndex.py")
source_python("src/dsmScript.py")
source_python('src/deephitter.py')
source_python("src/cIndex.py")
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)

epo=2
bsize=512
min_delta = 10^-7
patience = 1
iteration=5
########################
###sharedHyperparams
########################

layer1<-50
layer2<-50
layer3<-25
layer4<-25
drpt<-0.5

```

```{r ERSPCproofofconcept,eval=T}
##originally used 1 sample ,now using all data
data<-ERSPC
colnames(data)<-c("ScrArm","time","status")
  data<-as.data.frame(model.matrix(~.,data)[,-1])
mod_cb_glm <- fitSmoothHazard(status~.,
                              data =data,
                              time = "time",
                              event="status",
                              ratio = 100
)

covars_input<-layer_input(shape=c(length(colnames(mod_cb_glm$data))-2),
                          name = 'main_input')

covars_output<-covars_input%>% #layer_batch_normalization()%>%
  layer_dense(units=1,use_bias = T)

pmnn<-pmnnModel(features=colnames(mod_cb_glm$data)[-ncol(mod_cb_glm$data)],
                feature_input = covars_input,
                feature_output = covars_output,
                originalData = mod_cb_glm$data,
                offset=mod_cb_glm$offset,
                timeVar = "time",
                eventVar= "status",
                optimizer= 'adam'
)

fit<-fitSmoothHaz(pmnn,epochs=12000,#15000
                  batch_size = 1024,#159893,#
                  verbose=0,monitor = "loss",val_split=0)
times<-seq(0,max(data$time),length.out=100)
get_weights(fit$model)
coef(mod_cb_glm)

startOG <- Sys.time()
annPreds<-aarOld(fit,
              times=times,
              x_test=data[1:100,]
)
endOG <- Sys.time()
startq <- Sys.time()
annQuick<-aar(fit,times=times,x_test=data[1:100,])
endq <- Sys.time()

startOG-endOG
startq-endq

plot(annPreds,type="l")
lines(absoluteRisk(mod_cb_glm,time=times,newdata=data[1,],type="CI"))

results<-data.frame(intercept=c(coef(mod_cb_glm)[1],get_weights(fit$model)[[2]][1]),
                    ScrArm=c(coef(mod_cb_glm)[2],get_weights(fit$model)[[1]][1,1]),
                    time=c(coef(mod_cb_glm)[3],get_weights(fit$model)[[1]][2,1]))
rownames(results)<-c("Casebase","PMNN")
results<-round(results,2)

#saveRDS(results,"../paper/figures/proofOfConceptWeights.rds")

```

```{r}

sessionInfo()

```