### 2023-02-08 analyses/fullMetabricAnalysis.Rmd

---
title: "fullMetabricanalysis"
author: "Jesse"
date: "12/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r loadData}
library("illuminaHumanv4.db")
getExpressionDataM <- function(){
load("metabricData/Complete_METABRIC_Expression_Data.rbin")
expressionData<-Complete_METABRIC_Expression_Data@assayData$exprs

geneName<-data.frame(Gene=unlist(mget(x = rownames(expressionData),ifnotfound = NA,envir = illuminaHumanv4SYMBOL)))

geneName[which(is.na(geneName$Gene)),1]<-rownames(geneName)[which(is.na(geneName$Gene))]

rownames(expressionData)<-make.unique(geneName$Gene, sep="_")

return(t(expressionData))

}

load("metabricData/Complete_METABRIC_Clinical_Survival_Data_OS.rbin")

```

```{r processData}

library(survival)
timeEvent<-as.data.frame(Complete_METABRIC_Clinical_Survival_Data_OS)

timeEvent<-tidyr::separate(data = timeEvent, col =  "x",into =  c("time", "status"), sep = "\\+", extra = "merge")

timeEvent$status[which(is.na(timeEvent$status))]<-1

timeEvent$status[which(timeEvent$status=="")]<-0

#rm(Complete_METABRIC_Expression_Data)
rm(Complete_METABRIC_Clinical_Survival_Data_OS)
#rm(geneName)
gc()

```

```{r splitData}
set.seed(1)
data<-getExpressionDataM()

 spec = c(train = .7, test = .1, validate = .2)

 g = sample(cut(
   seq(nrow(data)), 
   nrow(data)*cumsum(c(0,spec)),
   labels = names(spec)
 ))

 res = list()
res$train<-  data[which(g=="train"),]
 res$validate<-  data[which(g=="validate"),] 
 res$test<-  data[which(g=="test"),]

saveRDS(res,"expressionSplit.rds")

```

```{r featureSelection}
res<-readRDS("expressionSplit.rds")$train
spearCor<-cor(res,method = c( "spearman"))
d <- as.dist(1-spearCor)
rm(spearCor,res)
gc()
hc <- hclust(d, method="complete")
#get distance
#get hclust
rm(d)
gc()

saveRDS(hc,"hierarchicalClusteringSpearmanDistance.rds")
```

```{r,fig.height=25,fig.width=25}

hc<-readRDS("hierarchicalClusteringSpearmanDistance.rds")
set.seed(1)

plot(hc)

clusts <- cutree(hc,k=1000)
i=1
chosen=data.frame(reps=rep("representative",1000))

for(i in 1:length(unique(clusts))){

chosen$reps[i]<-sample(names(which(clusts==i))[!grepl("^ILMN",names(which(clusts==i)))],1)

}

```

```{r}

expData<-getExpressionDataM()

importantGenePos<-which(colnames(expData) %in% chosen$reps)
importantGenes<-expData[,importantGenePos]

toPC<-expData[,-importantGenePos]

pcs <- prcomp(toPC, center = TRUE,scale. = TRUE)

```