### 2023-04-05 analyses/plotterUpdated.Rmd

---
title: "plottingScript"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(casebase) #core
library(keras) #core
source("src/functions.R") #core
library(data.table)#corecr
library(mltools)#corecr

library(survival)#compare
library(survminer)#compare
library(ggplot2)#compare
library(colorspace)#compare
library(DT)
library(splines)
library(reticulate)

library(simsurv)
library(flexsurv)
library(survivalmodels)
library(visreg)

#library(tidyverse)
library(riskRegression)
library(pseudo)

library(tensorflow)
library(ggrepel)

library(dplyr)
library(flextable)

library(cowplot)
library(matrixStats)
library(tableone)
library(gridExtra)
library(grid)
library(gtable)
#tf$config$threading$set_intra_op_parallelism_threads(1)
#tf$config$threading$set_inter_op_parallelism_threads(1)

#set.seed(2)
set.seed(1)
epo=2000
bsize=512
min_delta = 10^-7
patience = 10
iteration=100
e_sup=F
eval_firstSim<-F
numSamp<-2000

pal <- function(col, border = "light gray", ...)  {
  n <- length(col)
  plot(0, 0, type="n", xlim = c(0, 1), xlim=c(0.05,0.95),ylim = c(0, 1),
       axes = FALSE, xlab = "", ylab = "", ...)
  rect(0:(n-1)/n, 0, 1:n/n, 1, col = col, border = border)
}
statCalcSim<-function(metricSims){

  toRemove <- which(apply(metricSims, 1, function(x){any(is.na(x))})==T)
  if(length(toRemove)>0){
    metricSims<-metricSims[-c(toRemove),]
  }
  metricSummary<-data.frame(model=metricSims$model,times=metricSims$times)
  metricSummary<-cbind(metricSummary,sds=rowSds(as.matrix(metricSims[,-c(1,2)])),means=rowMeans(as.matrix(metricSims[,-c(1,2)])))

  ci<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){mean(x)+c(-1.96,1.96)*sd(x)/sqrt(length(x))})
  rownames(ci)<-c('cilower95','ciupper95')
  pi<-apply(as.matrix(metricSims[,-c(1,2)]), 1, function(x){quantile(x,c(.025, .975))})
  rownames(pi)<-c('pilower95','piupper95')

  metricSummary<-cbind(metricSummary,t(ci),t(pi))

  return(metricSummary)
}

plotPerformance<-function(metricSummary,xlab="Follow-up time",method='Metric',iterations='NA',ticks=5){

  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','cbnn','CBNN_Poly','optimal',
                                "Optimal"),
                        color=qualitative_hcl(9,palette='Dark 3'),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                   'DeepHit','CBNN','CBNN','Optimal',
                                   "Optimal"))
  allModels$color[1]="#000000"
  allModels$color[2]="#3cff4b"#"#7C4C74" #deephit
  allModels$color[3]="#1e90ff"#"#4363ff" #"#F9A602"#CBLR
  allModels$color[4]="#C0C000"#"#f032e6"#"#ff194B"
  allModels$color[5]="#ff8231" #deepsurv
  allModels$color[6]="#911eff"#"#911eff"
  allModels$color[7]="#911eff"#"#42d4f4"
  allModels$color[8]="#f00000"#"#ff32e6"
  allModels$color[9]="#f00000"
  allModels<- allModels[order(allModels$trueName),]
  i <- sapply(metricSummary, is.factor)
  metricSummary[i] <- lapply(metricSummary[i], as.character)
  metricSummary<-metricSummary[complete.cases(metricSummary), ]
  mainTitle<-""
  metricSummary$color<-NA
  for(i in 1:nrow(metricSummary)){
    modelPos<-which(metricSummary$model[i] ==allModels$model)
    metricSummary$model[i]<-allModels$trueName[modelPos]
    metricSummary$color[i]<-allModels$color[modelPos]
  }

  metricSummary<-metricSummary[order(metricSummary$model),]

  #colorsToUse<- which(allModels$trueName %in% metricSummary$model)

  if(method=="C-Index"){
    method=expression("C"["IPCW"])
  }   
  metricSummary#<- metricSummary[-which(metricSummary$model=="CBNN_Spline"),]
  #mainTitle<-paste(method,' : 95% P.I. over ', iterations, ' iterations',sep='')
  colnames(metricSummary)[1]<-"Models"
  a<-ggplot(data=metricSummary, aes(x=times,y=means,col=Models))+
    geom_line() +
    geom_ribbon(aes(ymin=cilower95,ymax=ciupper95,fill=Models),linetype=0,alpha=0.3) +
    ggtitle(mainTitle) + 
    xlab(xlab) +
    ylab(method)+
    theme_minimal()+
    scale_color_manual(values=unique(metricSummary$color))+
    scale_fill_manual(values=unique(metricSummary$color))+
    scale_x_continuous(n.breaks=6)
  # scale_color_manual(labels = metricSummary$model, values = metricSummary$color)

  return(a)
}

```

```{r getAll,eval=T}

#age, sex, race, number of comorbidities, presence of diabetes, presence of dementia, presence of cancer, mean arterial blood pressure, heart rate, respiration rate, temperature, white blood cell count, serum’s sodium, and serum’s creatinine

getPlots<-function(briers,cidxs,auc,dMax){

  BrierIPAList<-readRDS(briers)

  cidxList<-readRDS(cidxs)
  AUClist<-readRDS(auc)

  ipaSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  brierSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  ibsSims=data.frame(model=BrierIPAList[[1]]$model,times=BrierIPAList[[1]]$times*dMax)
  ibsTable<-as.data.frame(matrix(data=NA,nrow=100,ncol=length(unique(BrierIPAList[[1]]$model))))
  colnames(ibsTable)<-unique(BrierIPAList[[1]]$model)
  count=1
  for(sim in BrierIPAList){
    ipaSims<-cbind(ipaSims, sim$IPA)
    brierSims<-cbind(brierSims,sim$Brier)
    ibsSims<-cbind(ibsSims,sim$IBS)
    for(mod in colnames(ibsTable)){

      ibsTable[count,which(mod ==colnames(ibsTable))]<-sim$IBS[sim$model==mod & sim$time==max(sim$time)]

    }
    count<-count+1
  }
  ibsTable<-na.omit(ibsTable)

  aucSims=data.frame(model=AUClist[[1]]$model,times=AUClist[[1]]$times*dMax)
  for(sim in AUClist){
    aucSims<-cbind(aucSims,sim$AUC)

  }

  ipaSummary<-statCalcSim(ipaSims)
  brierSummary<-statCalcSim(brierSims)
  ibsSummary<-statCalcSim(ibsSims)
  aucSummary<-statCalcSim(aucSims)
  ipa<-plotPerformance(ipaSummary,method="IPA",iterations=iteration  )
  brier<-plotPerformance(brierSummary,method="BS",iterations=iteration )
  ibs<-plotPerformance(ibsSummary,method="IBS",iterations=iteration )
  auc<-plotPerformance(aucSummary,method="AUC",iterations=iteration )

  temp<-melt(setDT(as.data.frame(cidxList[[1]])), id.vars = c("times"), variable.name = "model")
  cidxSims=data.frame(model=temp$model,times=temp$times*dMax)

  count=1
  for(sim in cidxList){
    cidxSims<-cbind(cidxSims, melt(setDT(as.data.frame(sim)), id.vars = c("times"), variable.name = "model")$value)

    count<-count+1
  }

  cidxSummary<-statCalcSim(cidxSims)
  cidxC<-plotPerformance(cidxSummary,method="C-Index",iterations=iteration  )
  return(list(ipa=ipa,
              brier=brier,
              ibs=ibs,
              cidx=cidxC,
              auc=auc))
}

cMax<-max(readRDS("ComplexSimTest.rds")$time)/365
mMax<-max(readRDS("myelomaTest.rds")$time)/365
fMax<-max(readRDS("flchainTest.rds")$time)/365
pMax<-max(readRDS("prostateTest.rds")$time)/12

complex<-getPlots('results/100complexBrier.rds','results/100complexcidx.rds','results/100complexAuc.rds',cMax)

myeloma<-getPlots('results/100myelomaBrier.rds','results/100myelomacidx.rds','results/100myelomaAuc.rds',mMax)

flchain<-getPlots('results/100flchainBrier.rds','results/100flchaincidx.rds','results/100flchainAuc.rds',fMax)

prostate<-getPlots('results/100prostateBrier.rds','results/100prostatecidx.rds','results/100prostateAuc.rds',pMax)
#sup<-getPlots('results/100supportBrier.rds','results/100supportcidx.rds')

```

```{r megaplot,fig.height=12,fig.width=12}
#complex[[4]]$layers[[2]]<-NULL
pt <- plot_grid(
  complex[[1]]+ theme(legend.position="none")+ ggtitle("Complex simulation")+ labs(x = "")+coord_cartesian(ylim=c(-0.3,0.7)),
  myeloma[[1]]+ theme(legend.position="none")+ ggtitle("MM")+ labs(x = "",y="")+coord_cartesian(ylim=c(-1,.1)),
  flchain[[1]]+ theme(legend.position="none")+ ggtitle("FLC")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.025,.45)),
  prostate[[1]]+ theme(legend.position="none")+ ggtitle("Prostate")+ labs(x = "",y="")+coord_cartesian(ylim=c(-0.02,0.205)),
  complex[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.68,1)),
  myeloma[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.00,0.75)),
  flchain[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.8,0.875)),
  prostate[[5]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.82)),
  # complex[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)")+coord_cartesian(ylim=c(0.5,0.8)),
  # myeloma[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.4,0.65)),
  # flchain[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.65,0.9)),
  # prostate[[4]]+ theme(legend.position="none")+ labs(x = "Follow-up time (years)",y="")+coord_cartesian(ylim=c(0.5,0.73)),
  labels = c('A',"B","C","D",'E','F','G',"H"),#,"I","J","K","L"),
  ncol=4)
legend <- get_legend(
  # create some space to the left of the legend
  complex$ipa  +
    guides(color = guide_legend(nrow = 2)) +
    theme(legend.position = "bottom")+
    theme(legend.text=element_text(size=8))
)

g<-plot_grid(pt, legend,ncol = 1, rel_heights = c(1, .1), rel_widths = c(3, .4))
pdf(file="figures/megaPlot.pdf",width=16,height=9)
g
dev.off()
```

```{r flextableTablePrep,fig.height=46,fig.width=26}

edit_cell <- function(table, row, col, name="core-fg", ...){
  l <- table$layout
  ids <- which(l$t %in% row & l$l %in% col & l$name==name)
  for (id in ids){
    newgrob <- editGrob(table$grobs[id][[1]], ...)
    table$grobs[id][[1]] <- newgrob
  }
  table
}

#SUMMARY DATA ORDER, IPA BS IBS CIDX
makeTable<-function(summaryData,study="study"){
  allModels<-data.frame(model=c("Null model","Cox_Lin",'CB_Logi','DeepSurv',
                                'DeepHit','CBNN_Poly','Optimal'),
                        color=rainbow_hcl(7, start = 30, end = 300),
                        trueName=c("KM","Cox",'CBLR','DeepSurv',
                                   'DeepHit','CBNN','Optimal'))

  wantedScores<-c(0.25,0.5,0.75,1)

  tableResults<-as.data.frame(matrix(NA,nrow=nrow(allModels)+2,ncol=20))
  tableResults[1,]<-c("IPA","IPA","IPA","IPA",
                      "BS","BS","BS","BS",
                      "IBS","IBS","IBS","IBS",
                      "C-Index","C-Index","C-Index","C-Index",
                      "AUC","AUC","AUC","AUC"
  )
  tableResults[2,]<-c(0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1,0.25,0.5,0.75,1)*100
  indexer<-1
  rownames(tableResults)<-c(study,"% of time",allModels$trueName)
  meanOnly<-tableResults
  for (i in 1:length(summaryData)){ #IPA BRIER IBS CIDX is indexed by i

    tempPos<-which(rownames(tableResults) %in% unique(summaryData[[i]]$Models)  )
    tempModels<-rownames(tableResults)[ tempPos]
    for(j in 1:length(tempModels)){ #indexes model
      offset<-(i-1)*4
      indexer=1+offset
      for(k in 1:length(wantedScores)){ #indexes .25 .5 .75 1.0
        tempTime<-unique(summaryData[[i]]$times)[(floor(length(unique(summaryData[[i]]$times))*wantedScores[k]))]
        tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        if(length(tempResRow)==0){
          tempTime<-max(summaryData[[i]]$times[which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]])])
          tempResRow<-which(summaryData[[i]]$Models==rownames(tableResults)[ tempPos[j]] & summaryData[[i]]$times==tempTime)
        }
        # if(length(round(summaryData[[i]]$means[tempResRow],2))<1){
        #   tableResults[tempPos[j],indexer]<-"-0.23 (-0.25,-0.23)" #note that this is grabbed directly from the last estimate of Cox, as its the only time we don't get a value. So we extrapolate from the last available survival time.
        # }else{
        tableResults[tempPos[j],indexer]<-paste(round(summaryData[[i]]$means[tempResRow],2),
                                                " (",round(summaryData[[i]]$cilower95[tempResRow],2),
                                                ",",round(summaryData[[i]]$ciupper95[tempResRow],2),")",sep="")
        meanOnly[tempPos[j],indexer]<-round(summaryData[[i]]$means[tempResRow],2)
        #}
        indexer<-indexer+1

      }
    }

  }

  return(list(tab=tableResults[-3,-c(5:16)],means=meanOnly[-3,-c(5:16)]))#
}

cleantables<-function(tab){
  toRemove<-which(rowSums(is.na(tab))==ncol(tab))
  tab[is.na(tab)]<-"-"
  if(length(toRemove)>0){

    return(tab[-toRemove,])
  }
  return(tab)
}

quickEdit<-function(grob,rows,columns){
  return(edit_cell(grob, rows, columns, "core-fg", 
                   gp=gpar(fontsize=12, fontface="bold")))
}

addBorders<-function(grob){
  grob$layout[1:2 , c("l", "r")] <- list(c(5, 9),c(2,6))
  grob<- gtable_add_grob(grob,
                         grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                         t = 2, b = nrow(grob), l = 1, r = ncol(grob))
  grob <- gtable_add_grob(grob,
                          grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                          t = 1, l = 1, r = ncol(grob))
  grob <- gtable_add_grob(grob,
                          grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
                          t = 1, l = 2,r=5,b= nrow(grob))
  return(grob)
}

getMaxes<-function(Tab){

  tab<-Tab$tab

  sumer<-Tab$means[-c(1,2),]
  sumer<-as.data.frame(apply(na.omit(sumer),2,as.numeric))
  rowPick<-c()
  colPick<-c()

  rows<-apply(sumer,2,function(x){which(x==max(x))})
  cols<-c()
  for(i in 1:length(rows)){
    cols<-append(cols,rep(i,length(rows[[i]])))
  }
  rows<-c(5,as.vector(unlist(rows)))
  picks<-data.frame(row=rows,
                    col=c(0,cols))
  return(picks)
}

convert0<-function(tab){
  tempNames<-rownames(tab)
  tab<-as.data.frame(lapply(tab, function(x){gsub("0 ","0.00 ", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x){gsub('(0,',"(0.00,", x,fixed=T)}))
  tab<-as.data.frame(lapply(tab, function(x) gsub(",0)",",0.00)", x,fixed=T)))
  rownames(tab)<-tempNames
  return(tab)
}

flexMaker<-function(tab,boldPicks=NA,title){
  boldPicks$col=boldPicks$col+1 #makes finding the correct ones easier.
  bord<-fp_border_default(color="black",style="solid")
  bord2<-fp_border_default(color="black",style="solid",width = 2)
  tab<-cbind(rownames(tab),tab)
  ft<-delete_part(flextable(tab[-c(1,2),],cwidth=5,cheight=1), part = "header")

  ft <- add_header_row(x = ft, values = c(title,"IPA","AUC"),
                       colwidths = c(1,4,4))
  # ft <- compose(ft, part = "header", i = 1, j = 6,
  #               value = as_paragraph("C",
  #                                    as_sub("IPCW")))
  #ft<-bold(ft, bold = TRUE, part = "header")

  #ft$header$styles$text$bold$data
  ft <- add_header_row(x = ft,top=F, values = c("Method", 25,50,75,100,25,50,75,100),
                       colwidths = c(1,1,1,1,1,1,1,1,1))
  ft$header$styles$text$bold$data[1,1]<-T
  #ft <- add_header_row(
  #  x = ft, values = c(title),
  #  colwidths = c(9))
  ft <- align(ft, i = c(1,2), part = "header", align = "center")
  ft<-vline(ft, j = c(1,5), border = bord2, part = "body")
  ft<-hline(ft,i=c(1,2),border=bord2,part="header")
  #ft<-hline(ft,i=c(1),border=bord,part="body")
  ft<-fontsize(ft, i = NULL, j = NULL, size =7, part = "all")
  ft<-fontsize(ft, i = NULL, j = NULL, size =10, part = "header")
  ft <- align(ft, i = NULL, part = "body", align = "center")
  ft$body$styles$text$font.size$data[,1]<-10

  if(is.data.frame(boldPicks)){
    for(i in 1:nrow(boldPicks)){
      ft$body$styles$text$bold$data[boldPicks$row[i],boldPicks$col[i]]<-T
    }
  }
  return(rasterGrob(as_raster(autofit(ft))))
}

```

```{r flexTableMain,fig.height=46,fig.width=26}

getFlex<-function(result,titler="X: study"){
  Tab<-makeTable(list(result$ipa$data,result$brier$data,result$ibs$data,result$cidx$data,result$auc$data),study="Complex")
  Picks<-getMaxes(Tab)
  Tab<-cleantables(Tab$tab)
  Tab<-convert0(Tab)
  Flex<-flexMaker(tab=Tab,boldPicks = Picks,title=titler)

}

cFlex<-getFlex(complex,"A: Complex")
mFlex<-getFlex(myeloma,"B: MM")
fFlex<-getFlex(flchain,"C: FLC")
pFlex<-getFlex(prostate,"D: Prostate")

pdf(file="figures/megaTable.pdf",width=14,height=13)
cowplot::plot_grid(cFlex,mFlex,fFlex,pFlex,ncol=1)
dev.off()

```

```{r hyperparams}

getWinsTable<-function(win,main){
  winners<-readRDS(win)

  winners$cbnn$alp="-"
  winners$ds$alp="-"
  winners$ds<-winners$ds[,c(1,2,3,4,5,6,9,7,8)]

  winners$cbnn<-winners$cbnn[,c(1,2,3,4,5,6,9,7,8)]
  tabWin<-data.frame(CBNN=t(winners$cbnn),
                     DeepSurv=t(winners$ds),
                     DeepHit=t(winners$dh))

  colnames(tabWin)<-c("CBNN","DeepSurv","DeepHit")
  rownames(tabWin)<-c("Learning rate","Dropout","First layer nodes","Second layer nodes","Number of batches","Activation function","\U03B1","IBS","runtime")
  tabWin$`Hyper parameter`<-rownames(tabWin)
  tabWin<-tabWin[,c(4,1,2,3)]
  tabWin<-tabWin[-nrow(tabWin),]

  ft<-flextable(tabWin,cwidth=1.35,cheight=0.1)%>%flextable::hline(i = 7, part = 'body')
  ft <- add_header_row(x = ft, values = c(main),
                       colwidths = c(4))
  ft$header$styles$text$bold$data[1,1]<-T
  ft<-ft%>%hline_top(border = fp_border_default(width = 0), 
                     part = "header")

  return(rasterGrob(as_raster(autofit(ft))))
}

cwtab<-getWinsTable("complexOptimWins.rds","A: Complex")
mwtab<-getWinsTable("myelomaWins.rds","B: MM")
fwtab<-getWinsTable("flchainWins.rds","C: FLC")
pwtab<-getWinsTable("prostateWins.rds","D: Prostate")

pdf(file="figures/winTable.pdf",width=12,height=6)
cowplot::plot_grid(cwtab,mwtab,fwtab,pwtab,ncol=2)
dev.off()

```

```{r complexHazard}

library(casebase)

klist<-readRDS("ComplexSimCV.rds")
data<-rbind(klist[[1]]$train,klist[[1]]$val)

start_time <- Sys.time()
BrierIPAList<-list()
cScore<-list()
pos<-1

spec = c(train = .85, validate = .15)

g = sample(cut(
  seq(nrow(data)), 
  nrow(data)*cumsum(c(0,spec)),
  labels = names(spec)
))

res = split(data, g)

trainOG<-res$train

numSamp<-nrow(trainOG)
samp<-sample(seq(1,nrow(trainOG),by=1) , nrow(trainOG),replace = T) 
train<-trainOG[samp,]

#train<-res$train
val<-res$validate

sds<-sapply(train, function(x) (sd(x)))
means<-sapply(train, function(x) (mean(x)))
maxTime<-max(train$time)

fullTest<-readRDS("ComplexSimTest.rds")
test<-fullTest[,-c(ncol(data))]
test<-normalizer(fullTest,means,sds,maxTime)[,-c(ncol(data))]
fullTest<-normalizer(fullTest,means,sds,maxTime)

val<-normalizer(res$validate,means,sds,maxTime)

train<-normalizer(train,means,sds,maxTime)
linModelTrain<-as.data.frame(rbind(train,val))

fakeTrain<-train
fakeTrain$status[which((fakeTrain$status==1) & (fakeTrain$z1==max(fakeTrain$z1)))]<-2

dat<-popTime(fakeTrain,time="time",event="status")
a<-plot(dat, add.competing.event = T) +scale_color_manual(labels = c("Time-constant", "Time-varying"),values=c("red","blue"))+scale_fill_manual(labels = c("Time-constant", "Time-varying"),values=c("red","blue"))+
  guides(color = "none")+labs(fill="Case-series events")

pdf(file="figures/ComplexPopTime.pdf",width=8,height=6)
a
dev.off()

#https://onlinelibrary.wiley.com/doi/10.1111/biom.12232

```

```{r codeDump,eval=F,include=F}

times<-seq(from=min(test$time),
           to=max(test$time),
           length.out = 27
)

times<-head(times, -1)
times<-tail(times, -1)

haz_inter <- fitSmoothHazard(status~bs(time)+ z1+z2+z3+z1:time+z2:z3-time,
                             data = linModelTrain,
                             time = "time")

nonPH<-casebase::hazardPlot(haz_inter,newdata = data.frame(z1=max(train$z1),z2=1,z3=1))
nonPH$`Time Varying`<-TRUE
PH<- casebase::hazardPlot(haz_inter,newdata = data.frame(z1=min(train$z1),z2=0,z3=1))
PH$`Time Varying`<-FALSE

H<-as.data.frame(rbind(nonPH,PH))

ggplot(H,aes(x=time,y=predictedhazard,col=`Time Varying`))+geom_line()
fill_cols <- c("z1=0" = "Red", "z1=1" = "#009E73",
               "Base series" = "#0072B2")
g<-plot(popTime(train,time="time",event="status",),  add.competing.event=T,fill.params = list(
  name = element_blank(),
  breaks = c("Case series", "Competing event", "Base series")
))

g

a<-popTime(train,time="time",event="status",

)

plot(mod_cb_tvc,
     hazard.params = list(xvar = "time",
                          by = "z1",
                          alpha = 1,
                          ylab = "Hazard"))

mod_cb_tvc <- fitSmoothHazard(censrec ~ hormon * ns(log(rectime), df = 3),
                              data = brcancer,
                              time = "rectime")
plot(mod_cb_tvc,
     hazard.params = list(xvar = "rectime",
                          by = "hormon",
                          alpha = 0.05,
                          ylab = "Hazard"))

```